\LARGE\singlespacing{\textbf{S1 Appendix} \\ \medskip Estimating \emph{Mycobacterium} tuberculosis transmission in a South African clinic: Spatiotemporal model based on person movements}

### 2024-11-25 doc/paper/supplements.tex

\listoftables

@@

Person-tracking data required processing because many tracks were interrupted, for example, if the person i)~moved outside the range of the video sensors and then re-entered, ii)~bowed heavily downwards, or iii)~was temporarily hidden behind another person or obstacle. In addition, moving objects such as opening doors may have been recognised as a person, introducing many extremely short tracks (total duration of less than 10\,s and distance less than 2\,m), which were removed.

@@

The Shiny application is shown in \Cref{fig:shiny-app}. In the centre panel, a floor plan of the clinic is shown in (x,y)-space. Different areas of the clinic are delineated by colours, \eg seating areas are marked in purple and exit areas are marked in blue. Tracks are shown as lines with a open and crossed circles indicating the tracks' start and end, respectively. The currently selected track is shown in black, possible links in colours, and alternatives for the selected possible link in light grey.

Below the floor plan, a table is showing descriptive information about the possible links in relation to the currently selected track: (1)~the patient ID of the selected track (Pid) and the observation ID of possible links (Oid), (2)~the height difference at the end point of Pid and the start point of Oid, (3)~ the standing height (\ie maximum height during the entire track) difference, (4)~the duration of Oid, (5)~the time difference, and (6)~the distance between the end point of Pid and the start of point of Oid. Possible links are sorted in ascending order by the time difference.

The Shiny application's left panel shows descriptive information about the loaded file, the date of the tracking data, the initial number of IDs before any manual linkages, the current number of IDs after considering manual linkages, and the current number of made linkages. It further shows the current Pid with shortcuts to move to the next or previous Pid, or the next or previous Pid at the entrance. Linkages can be done either forward or backward in time. The total duration of the Pid track (time in the clinic) is shown together with the number of connected tracks within Pid. The sliding bars allow filtering of possible links (Oid) according to difference in time, distance, or height, with quick filters to switch between common thresholds. The first dropdown menu selects the Oid for which alternative matches should be shown in the floor plan. The idea is to prevent false linkages when there are more plausible alternatives for the Oid. A connection is made by selecting the Oid in the "Link with IDs" dropdown menu and then clicking "Link IDs". After that, the app updates the floor plan, matching table, and descriptive information. Erroneous link can be revoked by clicking "Unlink last ID", if the last link in time should be revoked, or by clicking "Unlink first ID", if the first link in time should be revoked. If the Pid could not be connected further, its track can be finished by entering a label and clicking "Label/End track". The person-tracking data is then updated with the reconnected track and saved automatically to the file shown at the bottom.

We created the following labels to denote finished tracks: 
\begin{itemize}
    \item \textbf{Clean}: A clean track throughout the clinic with known entrance and exit.
    \item \textbf{Staff}: IDs with moving in- and out of the staff room.
    \item \textbf{Noise}: A short track ($\leq$1\,min) without entrance or exit.
    \item \textbf{Lost enter}: A long track (>1\,min) without entrance.
    \item \textbf{Lost exit}: A long track (>1\,min) without exit.
    \item \textbf{Lost both}: A long track (>1\,min) without entrance and without exit.
\end{itemize}

@@

We obtained 3,319 tracks after preprocessing. Further exclusion of short tracks <5\,min resulted in 1,563 tracks used for analysis: 209 (13\%) labeled as "Clean", 125 (8\%) as "Staff", 213 (14\%) as "Lost enter", 246 (16\%) as "Lost exit", and 770 (49\%) as "Lost both".

@@

\section{Spatiotemporal modelling approach}\label{sec:spattemp-model}

\subsection{The Wells-Riley modelling framework}

We build upon the Wells-Riley model\cite{Riley1961Book}, which is an established epidemiological model to estimate the risk of infection. The model estimates the risk (probability) of airborne indoor transmission $P$ using the following equation: 
\begin{align}
    P = \frac{C}{S} = 1 - \exp \left(-\frac{Ipqt}{Q}\right),
\end{align}
where $C$ is the number of diseased cases, $S$ is the number of susceptible cases, $I$ is the number of infectious people in the indoor space, $p$ is the breathing rate (m$^3$\,h$^{-1}$), $q$ is the quantum (infectious dose) generation rate (quanta\,h$^{-1}$), $t$ is the exposure time (s), and $Q$ is the ventilation rate (m$^3$\,h$^{-1}$).

The unknown parameter $q$ is not the actual number of infectious particles in the air. It rather represents the dose of infectious particles that corresponds to a certain probability of infection with a Poisson relationship, \eg one quantum corresponds to $P = 1 - \exp (1) = 63\%$ risk of infection\cite{Rudnick2003IndoorAir}. It is an attempt to consider the stochastic behaviour of airborne transmission, since the number of infectious particles that cause an infection is unknown. Moreover, the risk of transmission depends on the environment, characteristics of the pathogen, and characteristics of the individual. For example, the type of respiratory activity (\eg breathing, coughing, or sneezing) determines the site in the respiratory tract where the exhaled particles are generated, which in turn influences the particle size distribution\cite{Wei2016AMJIC}. Smaller particles reach deeper into the lungs of the susceptible person\cite{Wang2021Science}, which increases the probability of infection.

While considering the stochastic behaviour of airborne transmission, the Wells-Riley model makes two simplifying assumptions. First, it assumes a well-mixed airspace, which means that the generated quanta disperse immediately and uniformly in the indoor airspace. In other words, the exposure to infectious quanta is the same regardless of the location of the infectious and susceptible individual. Second, the Wells-Riley model assumes steady-state conditions, which means that the quanta concentration and the outdoor air supply rate are constant over time. In other words, the risk of infection for each susceptible only varies with the duration of exposure (the time spent in the indoor space).

@@

The extension by Rudnick and Milton still assumes a well-mixed airspace. We relax this assumption by modelling spatial variation in the concentration of infectious quanta. We then compute the risk of infection for susceptible individuals using the Wells-Riley equation considering their cumulative exposure to infectious quanta over space and time as 
\begin{equation}\label{eq:spattemp-P}
    P = \sum_s \sum_t N_{c,t} \cdot \mathbb{I}_{c,t} \cdot p,
\end{equation}
where $N$ is the quanta concentration in the indoor space area $c$ at time $t$ (in quanta\,$m^3$), $\mathbb{I}$ is a binary variable indicating whether the susceptible individual was in $c$ at $t$, and $p$ is the breathing rate (in m$^3$\,h$^{-1}$).

Modelling of the spatiotemporal quanta concentration can be divided into three processes
\begin{itemize}
    \item[\ref{sec:quanta-generation}] \textbf{Quanta generation}: Infectious particles are generated at the infectious individual's location. 
    \item[\ref{sec:quanta-diffusion}] \textbf{Quanta diffusion}: Pathogen-carrying particles spread in the indoor space. 
    \item[\ref{sec:quanta-removal}] \textbf{Quanta removal}: Contaminated indoor air is replaced or pathogens become inactive. 
\end{itemize}
We describe each process in more detail in the following, but first we introduce some notation.

@@

We divide the airspace into a 2-dimensional grid with cells $c = 1, \dots, C$, where $(x_c, y_c)$ refer to the centre of cell $c$. The grid provides a discrete approximation to the continuously varying quanta concentration in the indoor space, with the number of grid cells $C = n_x \cdot n_y$ determining the spatial resolution. We denote the following variables and modelling parameters:
\begin{itemize}
    \item $N_{c,t}$: quanta concentration in airspace unit $c = 1, \dots, C$ at time $t = 1, \dots T$ (in seconds)
    \item $I_{c,t}$: number of infectious individuals in airspace unit $c$ at time $t$
    \item $q$: quanta generation rate (quanta s$^{-1}$)
    \item $D$: diffusion constant (m$^2$ s$^{-1}$)
    \item $AER$: air change rate (air changes s$^{-1}$)
    \item $V$: volume of the airspace (m$^3$)
    \item $CO^{\text{I}}$: CO$_2$ mixing ratio in indoor air (parts per million, ppm)
    \item $CO^{\text{O}}$: CO$_2$ mixing ratio in outdoor air (ppm)
    \item $G$: average CO$_2$ generation rate per person (L$\cdot$min$^{-1}\cdot$person$^{-1}$)
    \item $n_t$: number of individuals in the airspace at time $t$
    \item $\lambda$: bacterial inactivation rate (quanta s$^{-1}$)
    \item $k$: gravitational settling rate (quanta s$^{-1}$)
    \item $p$: breathing rate (m$^3$\,s$^{-1}$)
\end{itemize}

@@

The number of quanta generated by infectious individuals at time $t$ in cell $c$ is 
\begin{align}\label{eq:generation}
    I_{c,t} \cdot q ~.
\end{align}
We assume that the initially generated quanta is confined to the infectious individual's cell, reflecting previous findings that the concentration of pathogen-laden aerosols is typically highest near the infectious source\cite{Wang2021Science,Morawska2021,Vuorinen2020SafSci,Chen2020BuildEnv}. Although coughing and sneezing can expel the aerosols further away from the individual, these activities are much less frequent than breathing, which may thus contribute more infectious particles\cite{Dinkele2022AJRCCM}.

@@

The diffusion constant $D$ is unknown. We approximate it using an empirical relationship between the eddy diffusion coefficient $K$ ($K \approx D$ in m$^2$\,s$^{-1}$) and the outdoor air exchange rate $AER$ for CO$_2$ particles\cite{Cheng2011EnvSciTech,Foat2020BE}
\begin{align}
    K = (0.52 \cdot AER + 8.61\cdot10^{-5} \cdot \text{s}^{-1}) \cdot V^{\frac{2}{3}},
\end{align}
where $V$ is the volume of the airspace (in m$^3$). We thus assume that pathogen-carrying particles spread with a similar speed than CO$_2$ particles, which may overestimate the diffusion constant because CO$_2$ particles are smaller than \emph{Mtb}-carrying particles.

The $AER$ is not directly observed but can be estimated from indoor CO$_2$ levels, room occupancy, and volume. Under steady-state conditions (\ie CO$_2$ levels reaching a steady-state mixing ratio), the $AER$ can be computed as
\begin{align}
    AER = \frac{6\cdot10^4 \cdot n \cdot G}{V\cdot(CO^{\text{I}}-CO^{\text{O}})},
\end{align}
where $n$ is the number of people in the airspace, $G$ is the average CO$_2$ generation rate per person, $CO^{\text{I}}$ is the CO$_2$ mixing ratio in indoor and $CO^{\text{O}}$ is the mixing ratio in outdoor air\cite{Batterman2017IJERPH}. However, in a primary care clinic, room occupancy varies continuously and thus it may be difficult to determine the steady-state CO$_2$ mixing ratio. Therefore, we determine $AER$ using a transient mass balance model of the form\cite{Batterman2017IJERPH}
\begin{align}
    CO_{t+1}^{\text{I}} = \frac{6\cdot10^4 \cdot n_t + G}{Q} \cdot \left(1 - \exp(-Q/V \Delta t)\right) + (CO_t^{\text{I}}-CO^{\text{O}}) \cdot \exp(-Q/V \Delta t) + CO^{\text{O}},
\end{align}
where $Q$ is the outdoor air supply rate (in m$^3$\,s$^{-1}$ $\cdot \frac{1}{3600}$ ) and $AER = Q/V$. We numerically solve the transient mass balance model with the limited memory Broyden–Fletcher–Goldfarb–Shanno algorithm (L-BFGS), a quasi-Newton method, as implemented in the statistical software R\cite{Byrd1995SIAM}. The optimal solution is the outdoor air supply rate $Q$ that minimises the root mean square error. We also follow recommendations to fit the outdoor CO$_2$ mixing ratio with reasonable constraints on $CO^{\text{O}} = [300,600]$\,ppm\cite{Batterman2017IJERPH}. We can compute $AER$ over different time windows, hence the index in $AER_t$.

@@

Infectious quanta may be removed from the air through outdoor air exchange, bacterial inactivation, or gravitational settling. The removal rate is thus the sum of the outdoor air exchange rate $AER_t$, the bacterial inactivation rate $\lambda$ for \emph{Mtb}, and the gravitational settling rate $k$ for \emph{Mtb}-carrying particles, \ie 
\begin{align}\label{eq:removal}
    AER_t + \lambda ~.
\end{align}

@@

Combining \Cref{eq:generation}, \Cref{eq:diffusion}, and \Cref{eq:removal}, the quanta concentration at time $t$ in the indoor space can be computed as
\begin{align}\label{eq:spattemp-N}
    \underbrace{N_{c,t}}_{\text{new conc.}} = \underbrace{\left(D \Delta (\underbrace{N_{c,t-1}}_{\text{old conc.}} + \underbrace{I_{c,t} \cdot q}_{\text{generation}} \cdot V_{c}^{-1})\right)}_{\text{diffusion}} \cdot \underbrace{\exp\left(-(AER_{t} + \lambda + k\right)}_{\text{removal}},
\end{align}
where $V_c$ is the volume of cell $c$, assuming that the quanta concentration is vertically well-mixed.

@@

We illustrate our modelling approach with an example. Consider a room with a volume $V$=150\,m$^3$ (length = 10\,m, breadth = 5\,m, height = 3\,m), which we rasterise into a grid of $C = 40 \cdot 20 = 800$\,cells with an area of (0.25\,m)$^2$. We assume that quanta generation is confined to the cell where the infectious individual is located as well as the first neighbouring cells. We monitor the room from 8\,am to 12\,am and consider one infectious individual generating quanta continuously for one hour between 8\,am and 9\,am in the middle of the room (cell (20,10)), which means that quanta is generated in cells $(x_c = 20\pm1, y_c = 10\pm1)$. We assume a quanta generation rate of $q = 2$\,quanta h$^{-1}$ and a bacterial inactivation rate of $\lambda = 0$\,quanta h$^{-1}$. We further assume a constant outdoor air exchange rate of $AER = 1$\,air changes h$^{-1}$, translating into a quanta removal rate of 1\,quanta h$^{-1}$ and corresponding to a diffusion constant of 0.015\,m$^2$ s$^{-1}$.

\Cref{fig:toy-example} shows the spatiotemporal quanta concentration. The quanta concentration increases over time and is higher near the infectious individual in the room's centre (\Cref{fig:toy-example}a, top panel). After the individual has left the room, the quanta diffuses within minutes and most quanta is removed by 10\,am (\Cref{fig:toy-example}a, bottom panel) through outdoor air exchange. Spatial variation in the quanta concentration peaks at 9\,am, right before the infectious individual leaves the space, and then it takes about 15\,min until a well-mixed airspace has been reached through quanta diffusion (\Cref{fig:toy-example}b).

\begin{figure}[!htpb]
    \centering
    \includegraphics{tests/stm_v2-toy_example.png}
    \includegraphics{tests/stm_v2-toy_example-B.png}
    \caption[Example: Spatiotemporal quanta concentration in an indoor space]{\textbf{Example: Spatiotemporal quanta concentration in an indoor space between 8\,am and 10\,am following the generation of quanta by an infectious individual between 8\,am and 9\,am in the room's centre}. \textbf{(a)}~Spatial quanta concentration at selected time points and \textbf{(b)}~quanta concentration over time for the 800~cells into which the indoor space was divided.}
    \label{fig:toy-example}
\end{figure}

@@

We divide the clinic into three spaces: the waiting room (length = 10.55\,m $\times$ breadth = 5.50\,m $\times$ height = 3.00\,m), the corridor (length = 12.45\,m $\times$ breadth = 2.20\,m $\times$ height = 2.50\,m), and the TB room (length = 4.75\,m $\times$ breadth = 3.50\,m $\times$ height = 3.00\,m). Each room was discretised into a grid of cubic cells, with each cell covering a squared area of (0.25\,m)$^2$, corresponding to a diagonal cell length of $\sqrt{0.25^2 + 0.25^2} = 0.35$\,m. The volume of the cell depends on the room height, and we assume that the quanta concentration was vertically well-mixed.

Although the door between the waiting room and corridor was usually not closed, we consider these rooms as separate areas to facilitate modelling and because we observed differences in room occupancy and time-varying CO$_2$ levels. Nevertheless, we acknowledge the possibility for aerosolised \emph{Mtb} (quanta) to spread from the waiting room to the corridor, and vice versa, which is not considered in our simulations.

We model the spatiotemporal quanta concentration during clinic hours from 8\,am to 4\,pm and compare the modeled concentration by daytime, \ie morning (8\,am to 12\,am) and afternoon (12\,am to 4\,pm). Correspondingly, we estimate the outdoor air exchange rate by daytime, considering potential differences in ventilation conditions between the morning and afternoon. We update the quanta concentration every second from $t = 1, \dots, T$, in line with the frequency of the person-tracking data. We set the quanta concentration at $t=0$ to zero, \ie $N_{r,c,0} = 0 ~~ \forall r, c$. That is, we assume that all quanta from the previous day has been removed from the air before the start of the following clinic day.

@@

We use Monte Carlo simulation to estimate the risk of infection, considering uncertainty in several modelling parameters. Each simulation consists of the following steps:
\begin{itemize}
    \item[1.] \textbf{Sample uncertain modelling parameters}: Undiagnosed TB patients among clinic attendees, quanta generation rate, mask reduction rate, bacterial inactivation rate, and gravitational settling rate.
    \item[2.] \textbf{Compute quanta concentration}: Based on the generation of infectious quanta by diagnosed and undiagnosed TB patients, we compute the quanta concentration in each room and cell over time.
    \item[3.] \textbf{Estimate risk of infection}: Based on their spatiotemporal quanta exposure, we estimate the risk of infection for each clinic attendee using the Wells-Riley equation.
\end{itemize}

Many clinic attendees are potential undiagnosed TB patients and prior uncertainty in the quanta generation, mask reduction, bacterial inactivation, and gravitational settling rate is large. To reflect this uncertainty, we perform 5,000 Monte Carlo simulations and compute the individual risk of infection as the mean across these simulations.

\subsection{Modelling assumptions}\label{sec:priors}

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/modeling/mean-roi-assumptions.png}
    \caption[Mean risk of infection per clinic attendee when varying assumptions about the number of infectious people in the clinic]{Mean risk of infection per clinic attendee if \textbf{(a)}~the number of infectious people in the clinic corresponded to the prevalence of TB in the population, or \textbf{(b)}~the number of infectious people in the clinic corresponded to the number of bacteriologically confirmed TB patients and those suspected of having TB due to respiratory symptoms when attending the clinic. Low: <0.1\%, Medium: 0.1-1\%, High: >1\% risk of infection.}
    \label{fig:assumptions-results}
\end{figure}

>>>>

@@

Person-tracking data required processing because many tracks were interrupted, for example, if the person i)~moved outside the range of the video sensors and then re-entered, ii)~bowed heavily downwards, or iii)~was temporarily hidden behind another person or obstacle. In addition, moving objects such as opening doors may have been recognized as a person, introducing many extremely short tracks (total duration of less than 10\,s and distance less than 2\,m), which were removed.

@@

The Shiny application is shown in \Cref{fig:shiny-app}. In the center panel, a floor plan of the clinic is shown in (x,y)-space. Different areas of the clinic are delineated by colors, \eg seating areas are marked in purple and exit areas are marked in blue. Tracks are shown as lines with a open and crossed circles indicating the tracks' start and end, respectively. The currently selected track is shown in black, possible links in colors, and alternatives for the selected possible link in light gray.

Below the floor plan, a table is showing descriptive information about the possible links in relation to the currently selected track: (1)~the patient ID of the selected track (Pid) and the observation ID of possible links (Oid), (2)~the height difference at the end point of Pid and the start point of Oid, (3)~ the standing height (\ie maximum height during the entire track) difference, (4)~the duration of Oid, (5)~the time difference, and (6)~the distance between the end point of Pid and the start point of Oid. Possible links are sorted in ascending order by the time difference.

The Shiny application's left panel shows descriptive information about the loaded file, the date of the tracking data, the initial number of IDs before any manual linkages, the current number of IDs after considering manual linkages, and the current number of made linkages. It further shows the current Pid with shortcuts to move to the next or previous Pid, or the next or previous Pid at the entrance. Linkages can be done either forward or backward in time. The total duration of the Pid track (time in the clinic) is shown together with the number of connected tracks within Pid. The sliding bars allow filtering of possible links (Oid) according to difference in time, distance, or height, with quick filters to switch between common thresholds. The first dropdown menu selects the Oid for which alternative matches should be shown in the floor plan. The idea is to prevent false linkages when there are more plausible alternatives for the Oid. A connection is made by selecting the Oid in the "Link with IDs" dropdown menu and then clicking "Link IDs". After that, the app updates the floor plan, matching table, and descriptive information. Erroneous links can be revoked by clicking "Unlink last ID", if the last link in time should be revoked, or by clicking "Unlink first ID", if the first link in time should be revoked. If the Pid could not be connected further, its track can be finished by entering a label and clicking "Label/End track". The person-tracking data is then updated with the reconnected track and saved automatically to the file shown at the bottom.

We created the following labels to denote finished tracks: 
\begin{itemize}
    \item \textbf{Clean}: A clean track throughout the clinic with known entrance and exit.
    \item \textbf{Staff}: A track moving in- and out of the staff room.
    \item \textbf{Noise}: A short track ($\leq$1\,min) without entrance or exit.
    \item \textbf{Lost enter}: A long track (>1\,min) without entrance.
    \item \textbf{Lost exit}: A long track (>1\,min) without exit.
    \item \textbf{Lost both}: A long track (>1\,min) without entrance and without exit.
\end{itemize}

@@

We obtained 3,319 tracks after preprocessing. Further exclusion of relatively short tracks <5\,min resulted in 1,563 tracks used for analysis: 209 (13\%) labeled as "Clean", 125 (8\%) as "Staff", 213 (14\%) as "Lost enter", 246 (16\%) as "Lost exit", and 770 (49\%) as "Lost both".

@@

\section{Spatiotemporal modeling approach}\label{sec:spattemp-model}

\subsection{The Wells-Riley modeling framework}

We build upon the Wells-Riley model\cite{Riley1961Book}, which is an established epidemiological model to estimate the risk of infection. The model estimates the risk of airborne indoor transmission as
\begin{align}
    P = \frac{C}{S} = 1 - \exp \left(-\frac{Ipqt}{Q}\right),
\end{align}
where $P$ is the probability of infection, $C$ is the number of diseased cases, $S$ is the number of susceptible cases, $I$ is the number of infectious people in the indoor space, $p$ is the breathing rate (m$^3$\,h$^{-1}$), $q$ is the quantum (infectious dose) generation rate (quanta\,h$^{-1}$), $t$ is the exposure time (s), and $Q$ is the ventilation rate (m$^3$\,h$^{-1}$).

The unknown parameter $q$ is not the actual number of infectious particles in the air. It rather represents the dose of infectious particles that corresponds to a certain probability of infection with a Poisson relationship, \eg one quantum corresponds to $P = 1 - \exp (1) = 63\%$ risk of infection\cite{Rudnick2003IndoorAir}. It is an attempt to consider the stochastic behavior of airborne transmission, since the number of infectious particles that cause an infection is unknown. Moreover, the risk of transmission depends on the environment, characteristics of the pathogen, and characteristics of the individual. For example, the type of respiratory activity (\eg breathing, coughing, or sneezing) determines the site in the respiratory tract where the exhaled particles are generated, which in turn influences the particle size distribution\cite{Wei2016AMJIC}. Smaller particles reach deeper into the lungs of the susceptible person\cite{Wang2021Science}, which increases the probability of infection.

While considering the stochastic behavior of airborne transmission, the Wells-Riley model makes two simplifying assumptions. First, it assumes a well-mixed airspace, which means that the generated quanta disperse immediately and uniformly in the indoor airspace. In other words, the exposure to infectious quanta is the same regardless of the location of the infectious and susceptible individual. Second, the Wells-Riley model assumes steady-state conditions, which means that the quanta concentration and the outdoor air supply rate are constant over time. In other words, the risk of infection for each susceptible only varies with the duration of exposure (the time spent in the indoor space).

@@

The extension by Rudnick and Milton still assumes a well-mixed airspace. We relax this assumption by modeling spatial variation in the concentration of infectious quanta. We then compute the risk of infection for susceptible individuals using the Wells-Riley equation considering their cumulative exposure to infectious quanta over space and time as 
\begin{equation}\label{eq:spattemp-P}
    P = \sum_c \sum_t N_{c,t} \cdot \mathbb{I}_{c,t} \cdot p,
\end{equation}
where $N$ is the quanta concentration in a specific area of the indoor space $c$ at time $t$ (quanta\,m$^{-3}$), $\mathbb{I}$ indicates whether the individual was in $c$ at $t$, and $p$ is the breathing rate (m$^3$\,h$^{-1}$). Modeling of the spatiotemporal quanta concentration can subsequently be divided into three processes
\begin{itemize}
    \item[\ref{sec:quanta-generation}] \textbf{Quanta generation}: Infectious particles generated at the infectious individual's location. 
    \item[\ref{sec:quanta-diffusion}] \textbf{Quanta diffusion}: Pathogen-carrying particles disperse in the indoor space. 
    \item[\ref{sec:quanta-removal}] \textbf{Quanta removal}: Contaminated indoor air is replaced with fresh outdoor air. 
\end{itemize}
We describe each process in greater detail in the following, but first we introduce some notation.

@@

We assume that quanta concentration is well-mixed vertically and divide the airspace into a 2-dimensional grid with cells $c = 1, \dots, C$, where $(x_c, y_c)$ refer to the center of cell $c$. The grid provides a discrete approximation to the continuously varying quanta concentration in the indoor space, with the number of grid cells $C = n_x \cdot n_y$ determining the spatial resolution. We denote the following variables and modeling parameters:
\begin{itemize}
    \item $N_{c,t}$: quanta concentration in $c = 1, \dots, C$ at $t = 1, \dots T$ (in seconds)
    \item $I_{c,t}$: number of infectious individuals in $c$ at $t$
    \item $q$: quanta generation rate (quanta s$^{-1}$)
    \item $D$: diffusion constant (m$^2$ s$^{-1}$)
    \item $AER$: air change rate (s$^{-1}$)
    \item $V_c$: volume of the cell (m$^3$)
    \item $CO^{\text{I}}$: CO$_2$ mixing ratio in indoor air (parts per million, ppm)
    \item $CO^{\text{O}}$: CO$_2$ mixing ratio in outdoor air (ppm)
    \item $G$: average CO$_2$ generation rate per person (L$\cdot$min$^{-1}\cdot$person$^{-1}$)
    \item $n_t$: number of individuals in the airspace at time $t$
    \item $\lambda$: bacterial inactivation rate (s$^{-1}$)
    \item $k$: gravitational settling rate (s$^{-1}$)
    \item $p$: breathing rate (m$^3$\,s$^{-1}$)
\end{itemize}

@@

The number of quanta generated by infectious individuals at time $t$ in cell $c$ is 
\begin{align}\label{eq:generation}
    I_{c,t} \cdot q ~.
\end{align}
For now, we assume that the initially generated quanta is confined to the infectious individual's cell, reflecting previous findings that the concentration of pathogen-laden aerosols is typically highest near the infectious source\cite{Wang2021Science,Morawska2021,Vuorinen2020SafSci,Chen2020BuildEnv}. Although coughing and sneezing can expel the aerosols further away from the individual, these activities are much less frequent than breathing, which may thus contribute more infectious particles\cite{Dinkele2022AJRCCM}. Note, however, that the model can also accommodate assumptions about a larger initial spread of quanta.

@@

The diffusion constant $D$ is unknown. We approximate it using an empirical relationship between the eddy diffusion coefficient $K \approx D$ (m$^2$\,s$^{-1}$) and the outdoor air exchange rate $AER$ for CO$_2$ particles\cite{Cheng2011EnvSciTech,Foat2020BE}
\begin{align}
    K = (0.52 \cdot AER + 8.61\cdot10^{-5} \cdot \text{s}^{-1}) \cdot V^{\frac{2}{3}},
\end{align}
where $V$ is the volume of the entire space (m$^3$). Implicitly, we assume that pathogen-carrying particles spread with a similar speed than CO$_2$ particles, which may overestimate the diffusion constant because CO$_2$ particles are smaller than \emph{Mtb}-carrying particles.

The $AER$ is not directly observed but can be estimated from indoor CO$_2$ levels and room occupancy. Under steady-state conditions (\ie CO$_2$ levels reaching a steady-state mixing ratio), the $AER$ can be computed as
\begin{align}
    AER = \frac{6\cdot10^4 \cdot n \cdot G}{V\cdot(CO^{\text{I}}-CO^{\text{O}})},
\end{align}
where $n$ is the number of people in the airspace, $G$ is the average CO$_2$ generation rate per person, $CO^{\text{I}}$ is the CO$_2$ mixing ratio in indoor and $CO^{\text{O}}$ is the mixing ratio in outdoor air\cite{Batterman2017IJERPH}. However, in a primary care clinic, room occupancy varies continuously and thus it may be difficult to determine the steady-state CO$_2$ mixing ratio. Therefore, we determine $AER$ from time-varying CO$_2$ levels and room occupancy using a transient mass balance model of the form\cite{Batterman2017IJERPH}
\begin{align}
    CO_{t+1}^{\text{I}} = \frac{6\cdot10^4 \cdot n_t + G}{Q} \cdot \left(1 - \exp(-Q/V \Delta t)\right) + (CO_t^{\text{I}}-CO^{\text{O}}) \cdot \exp(-Q/V \Delta t) + CO^{\text{O}},
\end{align}
where $Q$ is the outdoor air supply rate (m$^3$\,s$^{-1}$ $\cdot \frac{1}{3600}$ ) and $AER = Q/V$. We numerically solve the transient mass balance model with the limited memory Broyden–Fletcher–Goldfarb–Shanno algorithm (L-BFGS), a quasi-Newton method, as implemented in the statistical software R\cite{Byrd1995SIAM}. The optimal solution is the outdoor air supply rate $Q$ that minimizes the root mean square error. We also follow recommendations to fit the outdoor CO$_2$ mixing ratio with reasonable constraints on $CO^{\text{O}} = [300,600]$\,ppm\cite{Batterman2017IJERPH}.

@@

Infectious quanta may be removed from the air through outdoor air exchange, bacterial inactivation, or gravitational settling. The removal rate is thus the sum of the outdoor air exchange rate $AER$, the bacterial inactivation rate $\lambda$ for \emph{Mtb}, and the gravitational settling rate $k$ for \emph{Mtb}-carrying particles, \ie 
\begin{align}\label{eq:removal}
    AER + \lambda + k~.
\end{align}

@@

Combining Equations \ref{eq:generation}, \ref{eq:diffusion}, and \ref{eq:removal}, the quanta concentration can be computed as
\begin{align}\label{eq:spattemp-N}
    \underbrace{N_{c,t}}_{\text{new conc.}} = \underbrace{\left(D \Delta (\underbrace{N_{c,t-1}}_{\text{old conc.}} + \underbrace{I_{c,t} \cdot q}_{\text{generation}} \cdot V_{c}^{-1})\right)}_{\text{diffusion}} \cdot \underbrace{\exp\left(-(AER_{t} + \lambda + k\right)}_{\text{removal}},
\end{align}
where $V_c$ is the volume of cell $c$.

@@

We illustrate our modeling approach with an example. Consider a room with a volume $V$=150\,m$^3$ (length = 10\,m, breadth = 5\,m, height = 3\,m), which we rasterize into a grid of $C = 40 \cdot 20 = 800$\,cells with an area of (0.25\,m)$^2$. We assume that quanta generation is confined to the cell where the infectious individual is located as well as the first neighboring cells. We monitor the room from 8\,am to 12\,am and consider one infectious individual generating quanta continuously for one hour between 8\,am and 9\,am in the middle of the room in cell (20,10), which means that quanta is generated in cells $(x_c = 20\pm1, y_c = 10\pm1)$. We assume a quanta generation rate of $q = 2$\,quanta h$^{-1}$ and a bacterial inactivation rate of $\lambda = 0$\,h$^{-1}$. We further assume a constant outdoor air exchange rate of $AER = 1$\,h$^{-1}$, corresponding to a diffusion constant of 0.015\,m$^2$ s$^{-1}$.

\Cref{fig:toy-example} shows the spatiotemporal quanta concentration. The quanta concentration increases over time and is higher near the infectious individual in the room's center (\Cref{fig:toy-example}a, top panel). After the individual has left the room, the quanta diffuses within minutes and most quanta is removed by 10\,am through outdoor air exchange (\Cref{fig:toy-example}a, bottom panel). Spatial variation in the quanta concentration peaks at 9\,am, right before the individual leaves the space, and 15\,min after that a well-mixed airspace has been reached through quanta diffusion (\Cref{fig:toy-example}b).

\begin{figure}[!htpb]
    \centering
    \includegraphics{tests/stm_v2-toy_example-A.png}
    \includegraphics{tests/stm_v2-toy_example-B.png}
    \caption[Example: Spatiotemporal quanta concentration in an indoor space]{\textbf{Example: Spatiotemporal quanta concentration in an indoor space between 8\,am and 10\,am following the generation of quanta by an infectious individual between 8\,am and 9\,am in the center of the room}. \textbf{(a)}~Spatial quanta concentration at selected time points and \textbf{(b)}~quanta concentration over time for the 800~cells into which the indoor space was rasterized.}
    \label{fig:toy-example}
\end{figure}

@@

We divide the clinic into three spaces: the waiting room (length = 10.55\,m $\times$ breadth = 5.50\,m $\times$ height = 3.00\,m), the corridor (length = 12.45\,m $\times$ breadth = 2.20\,m $\times$ height = 2.50\,m), and the TB room (length = 4.75\,m $\times$ breadth = 3.50\,m $\times$ height = 3.00\,m). Each room is discretized into a grid of cubic cells, with each cell covering a squared area of (0.25\,m)$^2$, corresponding to a diagonal cell length of $\sqrt{0.25^2 + 0.25^2} = 0.35$\,m. The volume of the cell depends on the room height.

Although the door between the waiting room and corridor was usually not closed, we consider these rooms to be separate to facilitate modeling. We acknowledge the possibility for aerosolized \emph{Mtb} to spread from the waiting room to the corridor, and vice versa, which is not considered.

We model the spatiotemporal quanta concentration during clinic hours from 8\,am to 4\,pm and compare the modeled concentration by daytime, \ie morning (8\,am to 12\,am) and afternoon (12\,am to 4\,pm). Correspondingly, we estimate the outdoor air exchange rate by daytime, considering potential differences in ventilation conditions between the morning and afternoon. We update the quanta concentration every second from $t = 1, \dots, T$, corresponding to the frequency of the person-tracking data. We set the quanta concentration at $t=0$ to zero, \ie $N_{r,c,0} = 0 ~~ \forall r, c$. That is, we assume that all quanta from the previous day has been removed from the air before the start of the following clinic day.

@@

We use Monte Carlo simulation to model the quanta concentration and estimate the individual risk of infection, considering uncertainty in several modeling parameters. Each simulation consists of the following steps:
\begin{itemize}
    \item[1.] \textbf{Sample uncertain modeling parameters}: Undiagnosed TB patients among clinic attendees, quanta generation rate, mask reduction rate, bacterial inactivation rate, and gravitational settling rate.
    \item[2.] \textbf{Model quanta concentration}: Based on the generation of infectious quanta by diagnosed and undiagnosed TB patients, we compute the quanta concentration in each room and cell over time.
    \item[3.] \textbf{Estimate risk of infection}: Based on their cumulative quanta exposure, we estimate the risk of infection for each clinic attendee using the Wells-Riley equation.
\end{itemize}

Many clinic attendees can be considered potential undiagnosed TB patients and prior uncertainty in the quanta generation, mask reduction, bacterial inactivation, and gravitational settling rate is large. To reflect this uncertainty, we perform 5,000 Monte Carlo simulations and compute the individual risk of infection as the mean across simulations.

\subsection{Modeling assumptions}\label{sec:priors}

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{results/modeling/mean-roi-assumptions.png}
    \caption[Mean risk of infection per clinic attendee when varying assumptions about the number of infectious people in the clinic]{Mean risk of infection per clinic attendee if \textbf{(a)}~the number of infectious people in the clinic corresponded to the prevalence of TB in the South African population, or \textbf{(b)}~the number of infectious people in the clinic corresponded to the number of bacteriologically confirmed TB patients and those suspected of having TB due to respiratory symptoms when attending the clinic. Low: <0.1\%, Medium: 0.1-1\%, High: >1\% risk of infection.}
    \label{fig:assumptions-results}
\end{figure}
